Structured Versus Decorated Cospans

### 2021-01-07 cospans.tex
tiny changes
small general edits

\begin{abstract}
One goal of applied category theory is to understand open systems.  We compare two ways of describing open systems as cospans equipped with extra data.    First, given a functor $L \maps \A \to \X$, a `structured cospan' is a diagram in $\A$ of the form $L(a) \rightarrow x \leftarrow L(b)$.  If $\A$ and $\X$ have finite colimits and $L$ preserves them, it is known that there is a symmetric monoidal double category whose objects are those of $\A$ and whose horizontal 1-cells are structured cospans.   Second, if $\A$ has finite colimits and $F \maps (\A,+) \to (\Cat,\times)$ is a symmetric lax monoidal pseudofunctor , a `decorated cospan' is a diagram in $\A$ of the form $a \rightarrow s \leftarrow b$ together with an object of $F(s)$.   Generalizing the work of Fong, we show that in this situation there is a symmetric monoidal double category whose objects are those of $\A$ and whose horizontal 1-cells are decorated cospans.   We prove that under certain conditions these two constructions become isomorphic when we take $\X = \inta F$ to be the Grothendieck category of $F$.  We illustrate these ideas with applications to electrical circuits, Petri nets, dynamical systems and epidemiological modeling.
\end{abstract}

@@

Here we give sufficient conditions for a decorated cospan double category to be equivalent---and in fact, isomorphic---to a structured cospan double category.  Suppose $\A$ has finite colimits and $F \maps (\A , +) \to (\Cat, \times)$ is a symmetric lax monoidal pseudofunctor.   Any such pseudofunctor gives an opfibration $U \maps \X \to \A$ where $\X = \inta F$ is defined by the Grothendieck construction.  Let $\Rex$ be the 2-category of categories with finite colimits, functors preserving finite colimits, and natural transformations.  We show that if $F$ factors through $\Rex$ as a pseudofunctor, the opfibration $U \maps \X \to \A$ is also a right adjoint.  From the accompanying left adjoint $L \maps \A \to \X$, we  construct a symmetric monoidal double category ${}_L \lCsp(\X)$ of structured cospans.  Finally, in \cref{thm:equiv} we prove that this structured cospan double category ${}_L \lCsp(\X)$ is isomorphic to the decorated cospan double category $F \lCsp$.   In fact, they are isomorphic as symmetric monoidal double categories.

@@

The category of objects $F\lCsp_0$ is simply $\A$, whereas $F$-decorated cospans and maps between them form the category of arrows $F\lCsp_1$. The (vertical) composition of two $F$-decorated cospan maps is 
\begin{equation}\label{eq:verticalcompo}
 \begin{tikzcd}
a\ar[r,"i"]\ar[d,"f"'] & m\ar[d,"h"] & b\ar[l,"o"']\ar[d,"g"]  & x\in F(m) \\
a'\ar[r,"{i'}"]\ar[d,"{f'}"'] & m'\ar[d,"{h'}"] & b'\ar[l,"{o'}"']\ar[d,"{g'}"] & x'\in F(m') \\
a''\ar[r,"{i''}"] & m'' & b''\ar[l,"{o''}"'] & x''\in F(m'')
 \end{tikzcd}\;=\;
 \begin{tikzcd}[column sep=.3in]
a\ar[r,"i"]\ar[dd,"{f'f}"'] & m\ar[dd,"{h'h}"] & b\ar[l,"o"']\ar[dd,"{g'g}"]  & x\in F(m) \\
\hole \\
a''\ar[r,"{i''}"] & m'' & b''\ar[l,"{o''}"'] & x''\in F(m'')  
 \end{tikzcd}
\end{equation}
where if $\tau \maps F(h)(x) \to x'$ and $\tau' \maps F(h')(x') \to x''$ are the decoration morphisms, 
\begin{equation}\label{eq:tauofverticalcomposite}
F(h'h)(x) \simrightarrow F(h')F(h)(x)\xrightarrow{F(h')(\tau)}F(h')(x')\xrightarrow{\tau'}x''
\end{equation}
is the decoration of the composite.

@@

where $\psi \maps m + n \to m+_{b} n$ is the natural map from the coproduct to the pushout and $\phi_{m,n} \maps F(m) \times F(n) \to F(m+n)$ are the lax monoidal structure maps of $F$. Moreover, given two horizontally composable maps of $F$-decorated cospans $\alpha$ and $\beta$ 

@@

\node () at (6.5,0) {$\cong$};%{$\stackrel{\phi_{h_1,h_2}}{\cong}$};

@@

\item of two 2-morphisms $\alpha$ and $\beta$ is:
\[
\begin{tikzpicture}[scale=1.2]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
%\node (E) at (2.7,0.5) {$\scriptstyle x \in F(m)$};
\node () at (2.7,0) {$\otimes$};
\node () at (6.2,0) {$=$};
%\node (F) at (2.7,-0.5) {$\scriptstyle x' \in F(m')$};
\node (G) at (3.5,0.5) {$c$};
\node (H) at (4.5,0.5) {$n$};
\node (I) at (5.5,0.5) {$d$};
\node (G') at (3.5,-0.5) {$c'$};
\node (H') at (4.5,-0.5) {$n'$};
\node (I') at (5.5,-0.5) {$d'$};
%\node (J) at (6.7,0.5) {$\scriptstyle y \in F(n)$};
%\node (K) at (6.7,-0.5) {$\scriptstyle y' \in F(n')$};
\node (L) at (1,-1.2) {$\scriptstyle\tau_\alpha \maps F(h)(x) \to x'\textrm{ in } F(m')$};
\node (M) at (4.5,-1.2) {$\scriptstyle\tau_\beta \maps F(h')(y) \to y'\textrm{ in } F(n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1$} (B)
(C) edge node[above]{$o_1$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i_1'$} (D)
(C') edge node[below] {$o_1'$} (D)
(B) edge node [left] {$h$} (D)
(G) edge node [above] {$i_2$} (H)
(G) edge node [left] {$f'$} (G')
(H) edge node [left] {$h'$} (H')
(G') edge node [below] {$i_2'$} (H')
(I) edge node [above] {$o_2$} (H)
(I) edge node [right] {$g'$} (I')
(I') edge node [below] {$o_2'$} (H');
\end{tikzpicture}
\begin{tikzpicture}[scale=1.1]
\node (A) at (0,0.5) {$a+c$};
\node (A') at (0,-0.5) {$a'+c'$};
\node (B) at (2,0.5) {$m+n$};
\node (C) at (4,0.5) {$b+d$};
\node (C') at (4,-0.5) {$b'+d'$};
\node (D) at (2,-0.5) {$m'+n'$};
%\node (E) at (5.2,0.5) {$\scriptstyle \phi_{m,n}(x,y)$};
%\node (F) at (5.2,-0.5) {$\scriptstyle \phi_{m',n'}(x',y')$};
\node () at (2,-1.3) {$\scriptstyle\tau_{\alpha\ot\beta}\maps F(h+h')(\phi_{m',n'}(x',y'))\to \phi_{m,n}(x,y)\textrm{ in }F(m'+n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1+i_2$} (B)
(C) edge node[above]{$o_1+o_2$} (B)
(A) edge node[left]{$f+f'$} (A')
(C) edge node[right]{$g+g'$} (C')
(A') edge node [below]{$i_1'+i_2'$} (D)
(C') edge node [below]{$o_1'+o_2'$} (D)
(B) edge node [left] {$h+h'$} (D);
\end{tikzpicture}
\]
with decoration morphism $\tau_{\alpha\otimes\beta}$ given by the following diagram:
\begin{displaymath}
 \begin{tikzcd}[column sep=.5in, row sep=.1in]
& F(m)\times F(n)\ar[r,"\phi_{m,n}"]\ar[dd,"F(h)\times F(h')"description]\ar[ddr,phantom,"\stackrel{\cref{eq:pseudonaturality}}{\cong}"] & F(m+n)\ar[dd,"F(h+h')"] \\
\one \ar[ur,"x\times y"]\ar[dr,"x'\times y'"']\ar[r,phantom,"\scriptstyle\Downarrow\tau_\alpha\times\tau_\beta"] & \phantom{A} && \\
& F(m')\times F(n')\ar[r,"\phi_{m',n'}"'] & F(m'+n')
 \end{tikzcd}
\end{displaymath}
\end{itemize}
\end{thm}

\begin{proof}
We first show that the categories of objects and arrows are symmetric monoidal, and that the source and target functors are symmetric strict monoidal.
Indeed, the category of objects $F\lCsp_0=\A$ is symmetric monoidal under the chosen binary coproducts in $\A$. The category of arrows $F\lCsp_1$ is also symmetric monoidal, with tensor product as in \cref{eq:tensordecoration} and monoidal unit
\[
I_1=\left(0\xrightarrow{!}0 \xleftarrow{!}0, \one \xrightarrow{\phi_0}F(0) \xrightarrow{F(!_0)} F(0) \right) 
\]
where $0$ is the initial object in $\A$.  By \cref{eq:UFCsp}, this unit $I_1$ is equal to $U_{0}$ as required.   In all that follows, we shall assume that $F(!_0)$ is the identity, so that
\begin{equation}\label{eq:monoidalunitFCsp}
I_1=\left(0\xrightarrow{!}0 \xleftarrow{!}0, \one \xrightarrow{\phi_0}F(0)\right).
\end{equation}
This can always be achieved by replacing $F$ by an equivalent pseudofunctor that strictly preserves identity morphisms \cite[Proposition 5.2]{LP}.  We do this merely to reduce the size of  our diagrams; the reader can reinsert the morphism $F(!_0)$ where needed.

The associator for the monoidal structure in $F\lCsp$ is formed as follows: for three objects
$M_1=(a_1\to m_1\leftarrow b_1,x_1\in F(m_1))$, $M_2=(a_2\to m_2\leftarrow b_2,x_2\in F(m_2))$ and $M_3=(a_3\to m_3\leftarrow b_3,x_3\in F(m_3),$
the isomorphism $(M_1\ot M_2)\ot M_3\cong M_1\ot(M_2\ot M_3)$ is the isomorphism of cospans
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$(a_1+a_2)+a_3$};
\node (A') at (0,-0.5) {$a_1+(a_2+a_3)$};
\node (B) at (2.25,0.5) {$(m_1+m_2)+m_3$};
\node (C) at (4.5,0.5) {$(b_1+b_2)+b_3$};
\node (C') at (4.5,-0.5) {$b_1+(b_2+b_3)$};
\node (D) at (2.25,-0.5) {$m_1+(m_2+m_3)$};
\node (E) at (7,0.5) {$(x_1 \otimes x_2) \otimes x_3 \in F((m_1{+}m_2){+}m_3)$};
\node (F) at (7,-0.5) {$x_1 \otimes (x_2 \otimes x_3)\in F(m_1{+}(m_2{+}m_3))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$(i_1{+}i_2){+}i_3$} (B)
(C) edge node[above]{$(o_1{+}o_2){+}o_3$} (B)
(A) edge node[left]{$\stackrel{\alpha}{\cong}$} (A')
(C) edge node[right]{$\stackrel{\alpha}{\cong}$} (C')
(A') edge node [below]{$i_1{+}(i_2{+}i_3)$} (D)
(C') edge node [below]{$o_1{+}(o_2{+}o_3)$} (D)
(B) edge node [left] {$\stackrel{\alpha}{\cong}$} (D);
\end{tikzpicture}
\]
together with the decoration isomorphism between $F(\alpha)((x_1 \otimes x_2) \otimes x_3)\stackrel{\cref{eq:tensordecoration}}{=}F(\alpha)(\phi_{m_1{+}m_2,m_3}(\phi_{m_1,m_2}(x_1,x_2),x_3))$ and $x_1 \otimes (x_2 \otimes x_3)=\phi_{m_1,m_2{+}m_3}(x_1,\phi_{m_2,m_3}(x_2,x_3))$ formed in the same way as the left two-piece part of \cref{eq:decoiso}. 
For the left and right unitors $I_1\ot M\cong M\cong M\ot I_1$ for any $M=(a\to m\leftarrow b,x\in F(m)$, we have span isomorphisms
\begin{displaymath}
 \begin{tikzcd}[column sep=.3in]
0+a\ar[rr,"1+i"]\ar[d,"\stackrel{\ell}{\cong}"'] && 0+m\ar[d,"\stackrel{\ell}{\cong}"'] && 0+b\ar[ll,"1+o"']\ar[d,"\stackrel{\ell}{\cong}"] &&1\ar[r,"\phi_0\times x"] & F(0)\times F(m)\ar[r,"\phi_{0,m}"] & F(0+m) \\
a\ar[rr,"i"] && m && b\ar[ll,"o"'] && \one \ar[r,"x"] & F(m)\\
a+0\ar[rr,"i+1"']\ar[u,"\stackrel{r}{\cong}"] && m+0\ar[u,"\stackrel{r}{\cong}"] && b+0\ar[ll,"o+1"]\ar[u,"\stackrel{r}{\cong}"']
&&\one \ar[r,"x\times\phi_0"] & F(m)\times F(0)\ar[r,"\phi_{m,0}"] & F(m+0)
 \end{tikzcd}
\end{displaymath}
with isomorphisms between decorations precisely given by \cref{eq:unitality}.
Finally, the braiding in $F\lCsp_1$ is inherited from that in $\A$ and $F$, with braiding $\beta\maps M_1\ot M_2\to M_2\ot M_1$ the cospan isomorphism
\begin{equation}\label{eq:braidingFCsp1}
 \begin{tikzcd}[column sep=.25in]
a_1+a_2\ar[rr,"i_1+i_2"]\ar[d,"\stackrel{\beta}{\cong}"'] && m_1+m_2\ar[d,"\stackrel{\beta}{\cong}"'] && b_1+b_2\ar[ll,"o_1+o_2"']\ar[d,"\stackrel{\beta}{\cong}"] &\one\ar[r,"x_1\times x_2"] & F(m_1)\times F(m_2)\ar[r,"\phi_{m_1,m_2}"] & F(m_1+m_2) \\
a_2+a_1\ar[rr,"i_2+i_1"'] && m_2+m_1 && b_2+b_1\ar[ll,"o_2+o_1"] &\one \ar[r,"x_2\times x_1"] & F(m_2)\times F(m_1)\ar[r,"\phi_{m_2,m_1}"] & F(m_2+m_1)
 \end{tikzcd}
\end{equation}
together with the decoration isomorphism 
\begin{equation}\label{eq:decobraiding}
 \begin{tikzcd}
  1\ar[r,"x_1\times x_2"]\ar[dr,"x_2\times x_1"'] & F(m_1)\times F(m_2)\ar[d,"\beta"']\ar[r,"\phi_{m_1,m_2}"]\ar[dr,phantom,"\stackrel{\cref{eq:braidedpseudofun}}{\cong}"] & F(m_1+m_2)\ar[d,"F(\beta)"] \\
  & F(m_2)\times F(m_1)\ar[r,"\phi_{m_2,m_1}"'] & F(m_2+m_1)
 \end{tikzcd}
\end{equation}
All axioms for a symmetric monoidal category can be verified, %see \cite[Theorem 4.1.3]{CourserThesis}, 
and it is easy to see that with these structures, the source and target functors $S,T\maps F\lCsp_1\to F\lCsp_0$ are symmetric strict monoidal.

@@

Similarly, for the comparison isomorphism between the units $U_{a+b}$ and $U_a\ot U_b$, it is easy to see that as cospans they are both the identity $1_{a+b} \maps a+b\to a+b\leftarrow a+b:1_{a+b}$, and there is a canonical isomorphism between their decorations 
\begin{displaymath}
\begin{tikzcd}[column sep=.55in]
\one\ar[r,"\phi_0"]\ar[dr,phantom,near start,"{\scriptstyle\stackrel{\cref{eq:unitality}}{\cong}}"]\ar[d,"\phi_0\times\phi_0"'] & F(0)\ar[dr,phantom,"{\scriptstyle\stackrel{\cref{eq:pseudonaturality}}{\cong}}"]\ar[r,"F(!_{a+b})"] & F(a+b)\ar[d,"F(1)"] \\
F(0)\times F(0)\ar[r,"F(!_a)\times F(!_n)"']\ar[ur,"\phi_{0,0}"description] & F(a)\times F(b)\ar[r,"\phi_{a,b}"'] & F(a+b)
\end{tikzcd}
\end{displaymath}
where the top composite is $\bot_{a+b}$ by \cref{eq:UFCsp} and the bottom composite is $\phi_{a,b}(\bot_a,\bot_b)$ by \cref{eq:tensordecoration}.
%, namely $F(!_{a+b})\phi_0\cong\phi_{a,b}(F(!_a)\times F(!_b))(\phi_0\times\phi_0)$ via the structure isomorphisms for $(\phi,\phi_0)$.

@@

Our construction gives not only decorated cospan double categories, but  also maps between these: that is, symmetric monoidal double functors, as in \cref{def:doublefun}. Suppose we have two categories $\A,A'$ with finite colimits and two symmetric lax monoidal pseudofunctors $F \maps \A \to \Cat$ and $F' \maps \A' \to \Cat$.    Then we can obtain a map between their decorated cospan double categories, $\lH \maps F\lCsp \to F' \lCsp$, from:

@@

In fact, all these correspondences lift to 2-equivalences.

>>>>

\begin{abstract}
One goal of applied category theory is to understand open systems.  We compare two ways of describing open systems as cospans equipped with extra data.    First, given a functor $L \maps \A \to \X$, a `structured cospan' is a diagram in $\A$ of the form $L(a) \rightarrow x \leftarrow L(b)$.  If $\A$ and $\X$ have finite colimits and $L$ preserves them, it is known that there is a symmetric monoidal double category whose objects are those of $\A$ and whose horizontal 1-cells are structured cospans. Second, given a pseudofunctor $F\maps\A\to\Cat$, a `decorated cospan' is a diagram in $\A$ of the form $a \rightarrow s \leftarrow b$ together with an object of $F(s)$. Generalizing the work of Fong, we show that if $\A$ has finite colimits and $F \maps (\A,+) \to (\Cat,\times)$ is symmetric lax monoidal, there is a symmetric monoidal double category whose objects are those of $\A$ and whose horizontal 1-cells are decorated cospans.
%Second, if $\A$ has finite colimits and $F \maps (\A,+) \to (\Cat,\times)$ is a symmetric lax monoidal pseudofunctor, a `decorated cospan' is a diagram in $\A$ of the form $a \rightarrow s \leftarrow b$ together with an object of $F(s)$.   Generalizing the work of Fong, we show that in this situation there is a symmetric monoidal double category whose objects are those of $\A$ and whose horizontal 1-cells are decorated cospans.   
We prove that under certain conditions, these two constructions become isomorphic when we take $\X = \inta F$ to be the Grothendieck category of $F$.  We illustrate these ideas with applications to electrical circuits, Petri nets, dynamical systems and epidemiological modeling.
\end{abstract}

@@

Here we give sufficient conditions for a decorated cospan double category to be equivalent---and in fact, isomorphic---to a structured cospan double category.  Suppose $\A$ has finite colimits and $F \maps (\A , +) \to (\Cat, \times)$ is a symmetric lax monoidal pseudofunctor.   Any such pseudofunctor gives an opfibration $U \maps \X \to \A$ where $\X = \inta F$ is defined by the Grothendieck construction.  Let $\Rex$ be the 2-category of categories with finite colimits, functors preserving finite colimits, and natural transformations.  We show that if $F$ factors through $\Rex$ as a pseudofunctor, the opfibration $U \maps \X \to \A$ is also a right adjoint. %%%
{\chris This is not a hundred percent true, because we want the corresponding F via the MonGroth to do so. This is the introduction though so we could be simpler, but more accurate perhaps?}
From the accompanying left adjoint $L \maps \A \to \X$, we  construct a symmetric monoidal double category ${}_L \lCsp(\X)$ of structured cospans.  Finally, in \cref{thm:equiv} we prove that this structured cospan double category ${}_L \lCsp(\X)$ is isomorphic to the decorated cospan double category $F \lCsp$.   In fact, they are isomorphic as symmetric monoidal double categories. %%%
{\chris Perhaps some ending statement saying something like ``this provides a passage between those two approaches, which we see in action in some of the main applications of the cospan frameworks'' or something? Right now it feels a bit like the paper finishes with theorem 3.2.}

@@

The category of objects $F\lCsp_0$ is simply $\A$, whereas $F$-decorated cospans and maps between them form the category of arrows $F\lCsp_1$. The (vertical) composition of two $F$-decorated cospan maps is 
\begin{equation}\label{eq:verticalcompo}
 \begin{tikzcd}
a\ar[r,"i"]\ar[d,"f"'] & m\ar[d,"h"] & b\ar[l,"o"']\ar[d,"g"]  & x\in F(m) \\
a'\ar[r,"{i'}"]\ar[d,"{f'}"'] & m'\ar[d,"{h'}"] & b'\ar[l,"{o'}"']\ar[d,"{g'}"] & x'\in F(m') \\
a''\ar[r,"{i''}"] & m'' & b''\ar[l,"{o''}"'] & x''\in F(m'')
 \end{tikzcd}\;=\;
 \begin{tikzcd}[column sep=.3in]
a\ar[r,"i"]\ar[dd,"{f'f}"'] & m\ar[dd,"{h'h}"] & b\ar[l,"o"']\ar[dd,"{g'g}"]  & x\in F(m) \\
\hole \\
a''\ar[r,"{i''}"] & m'' & b''\ar[l,"{o''}"'] & x''\in F(m'')  
 \end{tikzcd}
\end{equation}
where if $\tau \maps F(h)(x) \to x'$ and $\tau' \maps F(h')(x') \to x''$ are the decoration morphisms, 
\begin{equation}\label{eq:tauofverticalcomposite}
F(h'h)(x) \simrightarrow F(h')F(h)(x)\xrightarrow{F(h')(\tau)}F(h')(x')\xrightarrow{\tau'}x''
\end{equation}
is the decoration of the composite, where the first isomorphism comes from pseudofunctoriality of $F$.

@@

where $\psi \maps m + n \to m+_{b} n$ is the natural map from the coproduct to the pushout. Moreover, given two horizontally composable maps of $F$-decorated cospans $\alpha$ and $\beta$ 

@@

\node () at (6.5,0) {$\stackrel{\phi_{h_1,h_2}}{\cong}$};

@@

\item of two 2-morphisms $\alpha$ and $\beta$ is:
\[
\begin{tikzpicture}[scale=1.2]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
%\node (E) at (2.7,0.5) {$\scriptstyle x \in F(m)$};
\node () at (2.7,0) {$\otimes$};
\node () at (6.2,0) {$=$};
%\node (F) at (2.7,-0.5) {$\scriptstyle x' \in F(m')$};
\node (G) at (3.5,0.5) {$c$};
\node (H) at (4.5,0.5) {$n$};
\node (I) at (5.5,0.5) {$d$};
\node (G') at (3.5,-0.5) {$c'$};
\node (H') at (4.5,-0.5) {$n'$};
\node (I') at (5.5,-0.5) {$d'$};
%\node (J) at (6.7,0.5) {$\scriptstyle y \in F(n)$};
%\node (K) at (6.7,-0.5) {$\scriptstyle y' \in F(n')$};
\node (L) at (1,-1.2) {$\scriptstyle\tau_\alpha \maps F(h)(x) \to x'\textrm{ in } F(m')$};
\node (M) at (4.5,-1.2) {$\scriptstyle\tau_\beta \maps F(h')(y) \to y'\textrm{ in } F(n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1$} (B)
(C) edge node[above]{$o_1$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i_1'$} (D)
(C') edge node[below] {$o_1'$} (D)
(B) edge node [left] {$h$} (D)
(G) edge node [above] {$i_2$} (H)
(G) edge node [left] {$f'$} (G')
(H) edge node [left] {$h'$} (H')
(G') edge node [below] {$i_2'$} (H')
(I) edge node [above] {$o_2$} (H)
(I) edge node [right] {$g'$} (I')
(I') edge node [below] {$o_2'$} (H');
\end{tikzpicture}
\begin{tikzpicture}[scale=1.1]
\node (A) at (0,0.5) {$a+c$};
\node (A') at (0,-0.5) {$a'+c'$};
\node (B) at (2,0.5) {$m+n$};
\node (C) at (4,0.5) {$b+d$};
\node (C') at (4,-0.5) {$b'+d'$};
\node (D) at (2,-0.5) {$m'+n'$};
%\node (E) at (5.2,0.5) {$\scriptstyle \phi_{m,n}(x,y)$};
%\node (F) at (5.2,-0.5) {$\scriptstyle \phi_{m',n'}(x',y')$};
\node () at (2,-1.3) {$\scriptstyle\tau_{\alpha\ot\beta}\maps F(h+h')(\phi_{m',n'}(x',y'))\to \phi_{m,n}(x,y)\textrm{ in }F(m'+n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1+i_2$} (B)
(C) edge node[above]{$o_1+o_2$} (B)
(A) edge node[left]{$f+f'$} (A')
(C) edge node[right]{$g+g'$} (C')
(A') edge node [below]{$i_1'+i_2'$} (D)
(C') edge node [below]{$o_1'+o_2'$} (D)
(B) edge node [left] {$h+h'$} (D);
\end{tikzpicture}
\]
with decoration morphism $\tau_{\alpha\otimes\beta}$ given by the following diagram:
\begin{displaymath}
 \begin{tikzcd}[column sep=.5in, row sep=.1in]
& F(m)\times F(n)\ar[r,"\phi_{m,n}"]\ar[dd,"F(h)\times F(h')"description]\ar[ddr,phantom,"{\stackrel[\cref{eq:pseudonaturality}]{\phi_{h,h'}}{\cong}}"] & F(m+n)\ar[dd,"F(h+h')"] \\
\one \ar[ur,"x\times y"]\ar[dr,"x'\times y'"']\ar[r,phantom,"\scriptstyle\Downarrow\tau_\alpha\times\tau_\beta"] & \phantom{A} && \\
& F(m')\times F(n')\ar[r,"\phi_{m',n'}"'] & F(m'+n')
 \end{tikzcd}
\end{displaymath}
\end{itemize}
\end{thm}

\begin{proof}
We first show that the categories of objects and arrows are symmetric monoidal, and that the source and target functors are symmetric strict monoidal.
Indeed, the category of objects $F\lCsp_0=\A$ is symmetric monoidal under the chosen binary coproducts in $\A$. The category of arrows $F\lCsp_1$ is also symmetric monoidal, with tensor product as in \cref{eq:tensordecoration} and monoidal unit
\[
I_1=\left(0\xrightarrow{!}0 \xleftarrow{!}0, \one \xrightarrow{\phi_0}F(0) \xrightarrow{F(!_0)} F(0) \right) 
\]
where $0$ is the initial object in $\A$.  By \cref{eq:UFCsp}, this unit $I_1$ is equal to $U_{0}$ as required.   In all that follows, we shall assume that $F(!_0)$ is the identity, so that
\begin{equation}\label{eq:monoidalunitFCsp}
I_1=\left(0\xrightarrow{!}0 \xleftarrow{!}0, \one \xrightarrow{\phi_0}F(0)\right).
\end{equation}
This can always be achieved by replacing $F$ by an equivalent normal pseudofunctor, namely one that strictly preserves identity morphisms \cite[Proposition 5.2]{LP}.  We do this merely to reduce the size of  our diagrams; the reader can reinsert the morphism $F(!_0)$ where needed.

The associator for the monoidal structure in $F\lCsp$ is formed as follows: for three objects
$M_1=(a_1\to m_1\leftarrow b_1,x_1\in F(m_1))$, $M_2=(a_2\to m_2\leftarrow b_2,x_2\in F(m_2))$ and $M_3=(a_3\to m_3\leftarrow b_3,x_3\in F(m_3),$
the isomorphism $(M_1\ot M_2)\ot M_3\cong M_1\ot(M_2\ot M_3)$ is the isomorphism of cospans
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$(a_1+a_2)+a_3$};
\node (A') at (0,-0.5) {$a_1+(a_2+a_3)$};
\node (B) at (2.25,0.5) {$(m_1+m_2)+m_3$};
\node (C) at (4.5,0.5) {$(b_1+b_2)+b_3$};
\node (C') at (4.5,-0.5) {$b_1+(b_2+b_3)$};
\node (D) at (2.25,-0.5) {$m_1+(m_2+m_3)$};
\node (E) at (7,0.5) {$(x_1 \otimes x_2) \otimes x_3 \in F((m_1{+}m_2){+}m_3)$};
\node (F) at (7,-0.5) {$x_1 \otimes (x_2 \otimes x_3)\in F(m_1{+}(m_2{+}m_3))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$(i_1{+}i_2){+}i_3$} (B)
(C) edge node[above]{$(o_1{+}o_2){+}o_3$} (B)
(A) edge node[left]{$\stackrel{\alpha}{\cong}$} (A')
(C) edge node[right]{$\stackrel{\alpha}{\cong}$} (C')
(A') edge node [below]{$i_1{+}(i_2{+}i_3)$} (D)
(C') edge node [below]{$o_1{+}(o_2{+}o_3)$} (D)
(B) edge node [left] {$\stackrel{\alpha}{\cong}$} (D);
\end{tikzpicture}
\]
together with the decoration isomorphism between $F(\alpha)((x_1 \otimes x_2) \otimes x_3)\stackrel{\cref{eq:tensordecoration}}{=}F(\alpha)(\phi_{m_1{+}m_2,m_3}(\phi_{m_1,m_2}(x_1,x_2),x_3))$ and $x_1 \otimes (x_2 \otimes x_3)=\phi_{m_1,m_2{+}m_3}(x_1,\phi_{m_2,m_3}(x_2,x_3))$ formed in the same way as the left two-piece part of \cref{eq:decoiso}. 
For the left and right unitors $I_1\ot M\cong M\cong M\ot I_1$ for any $M=(a\to m\leftarrow b,x\in F(m)$, we have span isomorphisms
\begin{displaymath}
 \begin{tikzcd}[column sep=.3in]
0+a\ar[rr,"1+i"]\ar[d,"\stackrel{\ell}{\cong}"'] && 0+m\ar[d,"\stackrel{\ell}{\cong}"'] && 0+b\ar[ll,"1+o"']\ar[d,"\stackrel{\ell}{\cong}"] &&1\ar[r,"\phi_0\times x"] & F(0)\times F(m)\ar[r,"\phi_{0,m}"] & F(0+m) \\
a\ar[rr,"i"] && m && b\ar[ll,"o"'] && \one \ar[r,"x"] & F(m)\\
a+0\ar[rr,"i+1"']\ar[u,"\stackrel{r}{\cong}"] && m+0\ar[u,"\stackrel{r}{\cong}"] && b+0\ar[ll,"o+1"]\ar[u,"\stackrel{r}{\cong}"']
&&\one \ar[r,"x\times\phi_0"] & F(m)\times F(0)\ar[r,"\phi_{m,0}"] & F(m+0)
 \end{tikzcd}
\end{displaymath}
with isomorphisms between decorations precisely given by \cref{eq:unitality}.
Finally, the braiding in $F\lCsp_1$ is inherited from that in $\A$ and $F$, with braiding $\beta\maps M_1\ot M_2\to M_2\ot M_1$ the cospan isomorphism
\begin{equation}\label{eq:braidingFCsp1}
 \begin{tikzcd}[column sep=.25in]
a_1+a_2\ar[rr,"i_1+i_2"]\ar[d,"\stackrel{\beta}{\cong}"'] && m_1+m_2\ar[d,"\stackrel{\beta}{\cong}"'] && b_1+b_2\ar[ll,"o_1+o_2"']\ar[d,"\stackrel{\beta}{\cong}"] &\one\ar[r,"x_1\times x_2"] & F(m_1)\times F(m_2)\ar[r,"\phi_{m_1,m_2}"] & F(m_1+m_2) \\
a_2+a_1\ar[rr,"i_2+i_1"'] && m_2+m_1 && b_2+b_1\ar[ll,"o_2+o_1"] &\one \ar[r,"x_2\times x_1"] & F(m_2)\times F(m_1)\ar[r,"\phi_{m_2,m_1}"] & F(m_2+m_1)
 \end{tikzcd}
\end{equation}
together with the decoration isomorphism 
\begin{equation}\label{eq:decobraiding}
 \begin{tikzcd}
  1\ar[r,"x_1\times x_2"]\ar[dr,"x_2\times x_1"'] & F(m_1)\times F(m_2)\ar[d,"\beta"']\ar[r,"\phi_{m_1,m_2}"]\ar[dr,phantom,"{\stackrel[\cref{eq:braidedpseudofun}]{u_{a,b}}{\cong}}"] & F(m_1+m_2)\ar[d,"F(\beta)"] \\
  & F(m_2)\times F(m_1)\ar[r,"\phi_{m_2,m_1}"'] & F(m_2+m_1)
 \end{tikzcd}
\end{equation}
All axioms for a symmetric monoidal category can be verified,
and it is easy to see that with these structures, the source and target functors $S,T\maps F\lCsp_1\to F\lCsp_0$ are symmetric strict monoidal.

@@

Similarly, for the comparison isomorphism between the units $U_{a+b}$ and $U_a\ot U_b$, it is easy to see that as cospans they are both the identity $1_{a+b} \maps a+b\to a+b\leftarrow a+b:1_{a+b}$, and there is a canonical isomorphism between their decorations 
\begin{displaymath}
\begin{tikzcd}[column sep=.55in]
\one\ar[r,"\phi_0"]\ar[dr,phantom,near start,"{\scriptstyle\stackrel{\cref{eq:unitality}}{\cong}}"]\ar[d,"\phi_0\times\phi_0"'] & F(0)\ar[dr,phantom,"{\scriptstyle\stackrel{\cref{eq:pseudonaturality}}{\cong}}"]\ar[r,"F(!_{a+b})"] & F(a+b)\ar[d,"F(1)"] \\
F(0)\times F(0)\ar[r,"F(!_a)\times F(!_b)"']\ar[ur,"\phi_{0,0}"description] & F(a)\times F(b)\ar[r,"\phi_{a,b}"'] & F(a+b)
\end{tikzcd}
\end{displaymath}
where the top composite is $\bot_{a+b}$ by \cref{eq:UFCsp} and the bottom composite is $\phi_{a,b}(\bot_a,\bot_b)$ by \cref{eq:tensordecoration}.

@@

Our construction gives not only decorated cospan double categories, but  also maps between these: that is, symmetric monoidal double functors, as in \cref{def:doublefun}. Suppose we have two categories $\A,\A'$ with finite colimits and two symmetric lax monoidal pseudofunctors $F \maps \A \to \Cat$ and $F' \maps \A' \to \Cat$.    Then we can obtain a map between their decorated cospan double categories, $\lH \maps F\lCsp \to F' \lCsp$, from:

@@

In fact, all these correspondences lift to 2-equivalences between appropriate 2-categories.
